---
title: 'Weekly Exercises #3'
author: "Talia C"
output: 
  html_document:
    keep_md: TRUE
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_download: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)     
library(gardenR)       
library(lubridate)    
library(ggthemes)      
library(geofacet)      
theme_set(theme_minimal())       
```

```{r data}
# Lisa's garden data
data("garden_harvest")

# Seeds/plants (and other garden supply) costs
data("garden_spending")

# Planting dates and locations
data("garden_planting")

# Tidy Tuesday data
kids <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-15/kids.csv')
```


## Instructions

* Put your name at the top of the document. 

* **For ALL graphs, you should include appropriate labels.** 

* Feel free to change the default theme, which I currently have set to `theme_minimal()`. 

* Use good coding practice. Read the short sections on good code with [pipes](https://style.tidyverse.org/pipes.html) and [ggplot2](https://style.tidyverse.org/ggplot2.html). **This is part of your grade!**

* When you are finished with ALL the exercises, uncomment the options at the top so your document looks nicer. Don't do it before then, or else you might miss some important warnings and messages.



## Warm-up exercises with garden data

These exercises will reiterate what you learned in the "Expanding the data wrangling toolkit" tutorial. If you haven't gone through the tutorial yet, you should do that first.

  1. Summarize the `garden_harvest` data to find the total harvest weight in pounds for each vegetable and day of week (HINT: use the `wday()` function from `lubridate`). Display the results so that the vegetables are rows but the days of the week are columns.

```{r}
garden_harvest %>% 
  mutate(weight_lbs = (weight * 0.00220462), 
         day = wday(date, label = TRUE, abbr = TRUE)) %>% 
  group_by(vegetable, day) %>% 
  summarize(total_weight = sum(weight)) %>% 
  pivot_wider(id_cols = vegetable:day,
              names_from = "day", 
              values_from = "total_weight")
```


  2. Summarize the `garden_harvest` data to find the total harvest in pound for each vegetable variety and then try adding the plot from the `garden_planting` table. This will not turn out perfectly. What is the problem? How might you fix it?
  
The problem with this table is that there are multiple plots for the same vegetable and variety. This means that for example, the Bush Bush Slender variety of Beans has two different rows with the same total harvest weight in pounds but were harvested from different plots. The only way to fix this problem would be to combine the plots in some way so that for the beans the plot would say M and D. You could also maybe remove one of the plots but then you are removing a piece of the total harvest so the better way would be to combine them in a way.   


```{r}
garden_harvest %>% 
  group_by(vegetable, variety) %>% 
  mutate(weight_lbs = (weight * 0.00220462)) %>% 
  summarize(totharvest = sum(weight_lbs)) %>% 
  left_join(garden_planting, 
            by = c("vegetable", "variety")) %>% 
  summarise(vegetable, variety, totharvest, plot)
```


  3. I would like to understand how much money I "saved" by gardening, for each vegetable type. Describe how I could use the `garden_harvest` and `garden_spending` datasets, along with data from somewhere like [this](https://products.wholefoodsmarket.com/search?sort=relevance&store=10542) to answer this question. You can answer this in words, referencing various join functions. You don't need R code but could provide some if it's helpful.mdy
  
  In order to understand how much money you saved by gardening for each vegetable type, you would need to use a left join to the garden harvests data set to add on the money spent. This way you have the spending data in with the garden harvest data. Then you would need to create a data set with the amount of money it would cost to buy the vegetables and varieties from Whole Foods Market. After you do that, you will need to left join the whole foods data set to the new garden harvest data set. Then create a new variable using the mutate function that takes the price you paid for gardening and subtracts it from the price from whole foods. This will give you the amount of money you saved for each vegetable type. 



  4. Subset the data to tomatoes. Reorder the tomato varieties from smallest to largest first harvest date. Create a barplot of total harvest in pounds for each variety, in the new order.

```{r}
garden_harvest %>% 
  filter(vegetable == "tomatoes") %>% 
  group_by(variety) %>% 
  summarize(first_harvest = min(date)) %>% 
  arrange(first_harvest) 

#graph 
garden_harvest %>% 
  filter(vegetable == "tomatoes") %>% 
  group_by(variety) %>% 
  mutate(weight_lbs = weight * 0.00220462) %>% 
  summarize(first_harvest = min(date),
            total_harvest = sum(weight_lbs)) %>%
  arrange(first_harvest) %>% 
  ggplot(mapping = aes(x = total_harvest, 
                       y = fct_reorder(variety, first_harvest, .desc = TRUE))) +
  geom_col() +
  labs(x = NULL,
       y = NULL,
       title = "Total Harvest in Pounds for Each Variety",
       subtitle = "ordered from smallest to largest first date",
       caption = "Plot created by Talia C, data from gardenR package")
```


  5. In the `garden_harvest` data, create two new variables: one that makes the varieties lowercase and another that finds the length of the variety name. Arrange the data by vegetable and length of variety name (smallest to largest), with one row for each vegetable variety. HINT: use `str_to_lower()`, `str_length()`, and `distinct()`.
  
```{r}
garden_harvest %>%
  mutate(variety_lower = str_to_lower(variety),
         name_length = str_length(variety)) %>%
  distinct(vegetable, variety_lower, .keep_all = TRUE) %>% 
  arrange(vegetable, name_length)
```



  6. In the `garden_harvest` data, find all distinct vegetable varieties that have "er" or "ar" in their name. HINT: `str_detect()` with an "or" statement (use the | for "or") and `distinct()`.

```{r}
garden_harvest %>%
  mutate(variety_lower = str_to_lower(variety),
         name_length = str_length(variety),
         has_ar_er = str_detect(variety_lower, "ar | er")) %>%
  group_by(vegetable, variety_lower) %>% 
  distinct(vegetable, variety_lower, .keep_all = TRUE) 
```



## Bicycle-Use Patterns

In this activity, you'll examine some factors that may influence the use of bicycles in a bike-renting program.  The data come from Washington, DC and cover the last quarter of 2014.

Two data tables are available:

- `Trips` contains records of individual rentals
- `Stations` gives the locations of the bike rental stations

Here is the code to read in the data. We do this a little differently than usualy, which is why it is included here rather than at the top of this file. To avoid repeatedly re-reading the files, start the data import chunk with `{r cache = TRUE}` rather than the usual `{r}`.

```{r cache=TRUE}
data_site <- 
  "https://www.macalester.edu/~dshuman1/data/112/2014-Q4-Trips-History-Data.rds" 
Trips <- readRDS(gzcon(url(data_site)))
Stations<-read_csv("http://www.macalester.edu/~dshuman1/data/112/DC-Stations.csv")
```


### Temporal patterns

It's natural to expect that bikes are rented more at some times of day, some days of the week, some months of the year than others. The variable `sdate` gives the time (including the date) that the rental started. Make the following plots and interpret them:

  7. A density plot, which is a smoothed out histogram, of the events versus `sdate`. Use `geom_density()`.
  
```{r}
Trips %>% 
  ggplot(mapping = aes(x = sdate)) + 
  geom_density()+
  labs(x = NULL,
       y = NULL,
       title = "Density of Events Over the Time",
       caption = "Plot Created by Talia, data from Trips")
```
  
  8. A density plot of the events versus time of day.  You can use `mutate()` with `lubridate`'s  `hour()` and `minute()` functions to extract the hour of the day and minute within the hour from `sdate`. Hint: A minute is 1/60 of an hour, so create a variable where 3:30 is 3.5 and 3:45 is 3.75.
  
```{r}
Trips %>% 
  mutate(min_d = (minute(sdate))/60, 
         hour_d = hour(sdate), 
         hrmin = hour_d + min_d) %>% 
  summarize(hrmin, min_d, hour_d) %>% 
  ggplot(mapping = aes(x = hrmin)) +
  geom_density(fill = "light blue") +
  labs(x = NULL,
       y = NULL,
       title = "Density of Events Over the Time of Day",
       caption = "Plot Created by Talia, data from Trips")
```
  
  9. A bar graph of the events versus day of the week. Put day on the y-axis.
  
```{r}
Trips %>% 
  mutate(day = wday(sdate, label = TRUE)) %>%
  ggplot(mapping = aes(y = day)) +
  geom_bar() +
  labs(x = NULL, 
       y = NULL, 
       title = "Number of Events for Each Day of the Week",
       caption = "Plot Created by Talia, data from Trips")
```
  
  
  10. Facet your graph from exercise 8. by day of the week. Is there a pattern?
  
  The pattern is that on the weekdays, the amount distribution of events follows the normal work day schedule. Most events occur in the moring hours and in the late afternoon hours when people would be commuting to and from work. On the weekend, it is more during the mid day because people are not commuting to and from work and people are more likely to explore during the mid day over the weekend which means more bike usage by casual riders. 
  
```{r}
Trips %>% 
  mutate(day = wday(sdate, label = TRUE),
         min_d = (minute(sdate))/60, 
         hour_d = hour(sdate), 
         hrmin = hour_d + min_d) %>% 
  ggplot(mapping = aes(x = hrmin)) +
  geom_density() +
  facet_wrap(vars(day)) +
  labs(x = NULL, 
       y = NULL, 
       title = "Distribution of Events for Each Day of the Week",
       caption = "Plot Created by Talia, data from Trips")
```
  
  
The variable `client` describes whether the renter is a regular user (level `Registered`) or has not joined the bike-rental organization (`Causal`). The next set of exercises investigate whether these two different categories of users show different rental behavior and how `client` interacts with the patterns you found in the previous exercises. 

  11. Change the graph from exercise 10 to set the `fill` aesthetic for `geom_density()` to the `client` variable. You should also set `alpha = .5` for transparency and `color=NA` to suppress the outline of the density function.
  
```{r}
Trips %>% 
  mutate(day = wday(sdate, label = TRUE),
         min_d = (minute(sdate))/60, 
         hour_d = hour(sdate), 
         hrmin = hour_d + min_d) %>% 
  ggplot(mapping = aes(x = hrmin, fill = client)) +
  geom_density(alpha = 0.5, color = NA) +
  facet_wrap(vars(day)) +
  labs(x = NULL, 
       y = NULL,
       fill = "Client Level",
       title = "Distribution of Events for Each Day of the Week",
       caption = "Plot Created by Talia, data from Trips")
```

  12. Change the previous graph by adding the argument `position = position_stack()` to `geom_density()`. In your opinion, is this better or worse in terms of telling a story? What are the advantages/disadvantages of each?
  
  I think this is a worse in terms of telling a story because it is confusing to see the two stacked upon one another. The advantage of this graph is that one can clearly see the difference in the rider level and the distribution of events. However, the prior makes it easier to compare the two rider levels by seeing them on top of one another. 
  
```{r}
Trips %>% 
  mutate(day = wday(sdate, label = TRUE),
         min_d = (minute(sdate))/60, 
         hour_d = hour(sdate), 
         hrmin = hour_d + min_d) %>% 
  ggplot(mapping = 
           aes(x = hrmin, 
               fill = client)) +
  geom_density(alpha = 0.5, 
               color = NA,
               position = position_stack()) +
  facet_wrap(vars(day)) +
  labs(x = NULL, 
       y = NULL,
       fill = "Client Level",
       title = "Distribution of Events for Each Day of the Week",
       caption = "Plot Created by Talia, data from Trips")
```
  
  
  13. In this graph, go back to using the regular density plot (without `position = position_stack()`). Add a new variable to the dataset called `weekend` which will be "weekend" if the day is Saturday or Sunday and  "weekday" otherwise (HINT: use the `ifelse()` function and the `wday()` function from `lubridate`). Then, update the graph from the previous problem by faceting on the new `weekend` variable. 
  
```{r}
Trips %>% 
  mutate(day = wday(sdate, label = TRUE),
         min_d = (minute(sdate))/60, 
         hour_d = hour(sdate), 
         hrmin = hour_d + min_d,
         weekend = ifelse(day %in% c("Sat", "Sun"), "Weekend", "Weekday")) %>% 
  ggplot(mapping = 
           aes(x = hrmin, 
               fill = client)) +
  geom_density(alpha = 0.5, 
               color = NA) +
  facet_wrap(vars(weekend)) +
  labs(x = NULL, 
       y = NULL,
       fill = "Client Level",
       title = "Distribution of Events for Weekends and Weekdays",
       caption = "Plot Created by Talia, data from Trips")
```
  
  
  14. Change the graph from the previous problem to facet on `client` and fill with `weekday`. What information does this graph tell you that the previous didn't? Is one graph better than the other?
  
  This graph tells the information of distribution of when the events occurred either on the weekday or weekend by client level. While the prior graph communicates the distribution of the client level on a weekend or weekday. I wouldn't say one is better than the other but I think the prior graph communicates the data in a better way than this graph. 
  
```{r}
Trips %>% 
  mutate(day = wday(sdate, label = TRUE),
         min_d = (minute(sdate))/60, 
         hour_d = hour(sdate), 
         hrmin = hour_d + min_d,
         weekend = ifelse(day %in% c("Sat", "Sun"), "Weekend", "Weekday")) %>% 
  ggplot(mapping = 
           aes(x = hrmin, 
               fill = weekend)) +
  geom_density(alpha = 0.5, 
               color = NA) +
  facet_wrap(vars(client)) +
  labs(x = NULL, 
       y = NULL,
       fill = "",
       title = "Distribution of Events by Client Level",
       caption = "Plot Created by Talia, data from Trips")
```



### Spatial patterns

  15. Use the latitude and longitude variables in `Stations` to make a visualization of the total number of departures from each station in the `Trips` data. Use either color or size to show the variation in number of departures. We will improve this plot next week when we learn about maps!
  
```{r}
Trips %>% 
  group_by(sstation) %>% 
  summarize(total_departures = n()) %>% 
  left_join(Stations, 
            by = c("sstation" = "name")) %>% 
  ggplot(mapping = aes(x = long, 
                       y = lat,
                       color = total_departures)) +
  geom_point() +
  labs(x = NULL, 
       y = NULL, 
       title = "Total Departures from Different Stations",
       caption = "Plot created by Talia, data from Trips")
```
  
  
  16. Only 14.4% of the trips in our data are carried out by casual users. Create a plot that shows which area(s) have stations with a much higher percentage of departures by casual users. What patterns do you notice? (Again, we'll improve this next week when we learn about maps).
  
  I notice that the there is a higher proportion of casual riders on towards the center of the cluster of points and also on the outlier cluster of points.  
  
```{r}
Trips %>% 
  group_by(sstation) %>% 
  summarize(total_departures = n(),
            prop_casual = mean(client == "Casual")) %>% 
  left_join(Stations, 
            by = c("sstation" = "name")) %>% 
  ggplot(mapping = aes(x = long, 
                       y = lat,
                       color = prop_casual)) +
  geom_point() +
  labs(x = NULL, 
       y = NULL, 
       title = "Proportion of Casual Riders from Different Stations",
       caption = "Plot created by Talia, data from Trips")
```
  
### Spatiotemporal patterns

  17. Make a table with the ten station-date combinations (e.g., 14th & V St., 2014-10-14) with the highest number of departures, sorted from most departures to fewest. Save this to a new dataset and print out the dataset. Hint: `as_date(sdate)` converts `sdate` from date-time format to date format. 
  
```{r}
new_data_set <- Trips %>%
  mutate(justdate = as_date(sdate)) %>%
  group_by(sstation, justdate) %>% 
  summarize(total_departures = n()) %>% 
  arrange(desc(total_departures)) %>%
  distinct(sstation, justdate, .keep_all = TRUE) %>% 
  head(10) 
  
 head(new_data_set, 10)
```
  
  
  18. Use a join operation to make a table with only those trips whose departures match those top ten station-date combinations from the previous part.
  
```{r}
Trips %>% 
  mutate(justdate = as_date(sdate)) %>% 
  inner_join(new_data_set, by = c("justdate" = "justdate",
                                  "sstation" = "sstation")) %>% 
  group_by(sstation, justdate) %>% 
  summarize(total_departures = n()) %>% 
  arrange(desc(total_departures)) %>% 
  head(10)
```
  
  
  19. Build on the code from the previous problem (ie. copy that code below and then %>% into the next step.) and group the trips by client type and day of the week (use the name, not the number). Find the proportion of trips by day within each client type (ie. the proportions for all 7 days within each client type add up to 1). Display your results so day of week is a column and there is a column for each client type. Interpret your results.
  
  From these results, one can see that the proportion of casual riders greatly increases during the weekends and decreases during the weekdays. This makes sense because the registered riders are likely those who have to commute to and from work and so they use the stations more on the weekdays than the weekend.

```{r}
Trips %>% 
  mutate(justdate = as_date(sdate),
         day_ofweek = wday(sdate, label = TRUE)) %>% 
  inner_join(new_data_set, by = c("justdate" = "justdate",
                                  "sstation" = "sstation")) %>% 
  group_by(day_ofweek, client) %>% 
  count() %>% 
  group_by(client) %>% 
  mutate(prop = n/sum(n)) %>% 
  pivot_wider(id_cols = day_ofweek,
              names_from = client,
              values_from = prop)
```


## GitHub link

  20. Below, provide a link to your GitHub page with this set of Weekly Exercises. Specifically, if the name of the file is 03_exercises.Rmd, provide a link to the 03_exercises.md file, which is the one that will be most readable on GitHub.
  
  [Git hub link](https://github.com/tch8/weeklyexercise_3/blob/main/weekly-exercise--3-TC.md) 
  
  [git hub general link](https://github.com/tch8/weeklyexercise_3)


## Challenge problem! 

This problem uses the data from the Tidy Tuesday competition this week, `kids`. If you need to refresh your memory on the data, read about it [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-09-15/readme.md). 

  21. In this exercise, you are going to try to replicate the graph below, created by Georgios Karamanis. I'm sure you can find the exact code on GitHub somewhere, but **DON'T DO THAT!** You will only be graded for putting an effort into this problem. So, give it a try and see how far you can get without doing too much googling. HINT: use `facet_geo()`. The graphic won't load below since it came from a location on my computer. So, you'll have to reference the original html on the moodle page to see it.
  
```{r}
kids %>% 
  facet_geo(vars(state))
```