---
title: 'Weekly Exercises #4'
author: "Talia C"
output: 
  html_document:
    keep_md: TRUE
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_download: true
---


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)     
library(lubridate)     
library(openintro)     
library(palmerpenguins)
library(maps)          
library(ggmap)         
library(gplots)       
library(RColorBrewer)  
library(sf)            
library(leaflet)       
library(carData)       
library(ggthemes)      
theme_set(theme_minimal())
```

```{r data}
# Starbucks locations
Starbucks <- read_csv("https://www.macalester.edu/~ajohns24/Data/Starbucks.csv")

starbucks_us_by_state <- Starbucks %>% 
  filter(Country == "US") %>% 
  count(`State/Province`) %>% 
  mutate(state_name = str_to_lower(abbr2state(`State/Province`))) 

# Lisa's favorite St. Paul places - example for you to create your own data
favorite_stp_by_lisa <- tibble(
  place = c("Home", "Macalester College", "Adams Spanish Immersion", 
            "Spirit Gymnastics", "Bama & Bapa", "Now Bikes",
            "Dance Spectrum", "Pizza Luce", "Brunson's"),
  long = c(-93.1405743, -93.1712321, -93.1451796, 
           -93.1650563, -93.1542883, -93.1696608, 
           -93.1393172, -93.1524256, -93.0753863),
  lat = c(44.950576, 44.9378965, 44.9237914,
          44.9654609, 44.9295072, 44.9436813, 
          44.9399922, 44.9468848, 44.9700727)
  )

#COVID-19 data from the New York Times
covid19 <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")

```

## Instructions

* Put your name at the top of the document. 

* **For ALL graphs, you should include appropriate labels.** 

* Feel free to change the default theme, which I currently have set to `theme_minimal()`. 

* Use good coding practice. Read the short sections on good code with [pipes](https://style.tidyverse.org/pipes.html) and [ggplot2](https://style.tidyverse.org/ggplot2.html). **This is part of your grade!**

* When you are finished with ALL the exercises, uncomment the options at the top so your document looks nicer. Don't do it before then, or else you might miss some important warnings and messages.



## Warm-up exercises from tutorial

These exercises will reiterate what you learned in the "Mapping data with R" tutorial. If you haven't gone through the tutorial yet, you should do that first.

### Starbucks locations (`ggmap`)

  1. Add the `Starbucks` locations to a world map. Add an aesthetic to the world map that sets the color of the points according to the ownership type. What, if anything, can you deduce from this visualization?  
  
  I can see that the majority of Starbucks around the world are either company owned or licensed. I can also see that the density of Starbucks locations/number of locations is greatest in the USA specifically in the west and east side of the US. 
```{r}
#set map background 
world <- get_stamenmap(
    bbox = c(left = -180, bottom = -57, right = 179, top = 82.1), 
    maptype = "terrain",
    zoom = 2)

#plot points ontop of map 
ggmap(world) +
  geom_point(data = Starbucks, 
             aes(x = Longitude, y = Latitude, color = `Ownership Type`), 
             alpha = 0.5, 
             size = .3) +
  labs(title = "Ownership Type of Starbucks Locations around the World",
       caption = "Plot created by Talia C, data from Starbucks Data Set") +
  theme(plot.title.position = "plot")
```

  2. Construct a new map of Starbucks locations in the Twin Cities metro area (approximately the 5 county metro area).  
  
```{r}
#set Twin Cities map background 
TCmap <- get_stamenmap(
    bbox = c(left = -93.8605, bottom = 44.6111, right = -92.7207, top = 45.2267), 
    maptype = "terrain",
    zoom = 10)

#plot points ontop of map 
ggmap(TCmap) +
  geom_point(data = Starbucks, 
             aes(x = Longitude, y = Latitude, color = `Ownership Type`), 
             alpha = 1, 
             size = 1.5) + 
  labs(title = "Ownership Type of Starbucks Locations in the Twin Cities Metro Area",
       caption = "Plot created by Talia C, data from Starbucks Data Set") +
  theme(plot.title.position = "plot")
```

  3. In the Twin Cities plot, play with the zoom number. What does it do?  (just describe what it does - don't actually include more than one map).  
      When the zoom number increases to more than 10, the more zoomed out the map is and therefore when the zoom number decreases to less than 10, the map becomes more zoomed in. 

  4. Try a couple different map types (see `get_stamenmap()` in help and look at `maptype`). Include a map with one of the other map types.  

```{r}
#set Twin Cities map background 
TCmap2 <- get_stamenmap(
    bbox = c(left = -93.8605, bottom = 44.6111, right = -92.7207, top = 45.2267), 
    maptype = "toner",
    zoom = 10)

#plot points ontop of map 
ggmap(TCmap2) +
  geom_point(data = Starbucks, 
             aes(x = Longitude, y = Latitude, color = `Ownership Type`), 
             alpha = 1, 
             size = 1.5) +
  labs(title = "Ownership Type of Starbucks Locations in the Twin Cities Metro Area",
       caption = "Plot created by Talia C, data from Starbucks Data Set") +
  theme(plot.title.position = "plot")
```
  
  5. Add a point to the map that indicates Macalester College and label it appropriately. There are many ways you can do think, but I think it's easiest with the `annotate()` function (see `ggplot2` cheatsheet).

```{r}
#set Twin Cities map background 
TCmap <- get_stamenmap(
    bbox = c(left = -93.8605, bottom = 44.6111, right = -92.7207, top = 45.2267), 
    maptype = "terrain",
    zoom = 10)

#plot points ontop of map 
ggmap(TCmap) +
  geom_point(data = Starbucks, 
             aes(x = Longitude, y = Latitude, color = `Ownership Type`), 
             alpha = 1, 
             size = 1.5) +
  annotate("text", x = -93.1693, y = 44.9378, label = "Macalester College")
```





### Choropleth maps with Starbucks data (`geom_map()`)

The example I showed in the tutorial did not account for population of each state in the map. In the code below, a new variable is created, `starbucks_per_10000`, that gives the number of Starbucks per 10,000 people. It is in the `starbucks_with_2018_pop_est` dataset.

```{r}
census_pop_est_2018 <- read_csv("https://www.dropbox.com/s/6txwv3b4ng7pepe/us_census_2018_state_pop_est.csv?dl=1") %>% 
  separate(state, into = c("dot","state"), extra = "merge") %>% 
  select(-dot) %>% 
  mutate(state = str_to_lower(state))

starbucks_with_2018_pop_est <-
  starbucks_us_by_state %>% 
  left_join(census_pop_est_2018,
            by = c("state_name" = "state")) %>% 
  mutate(starbucks_per_10000 = (n/est_pop_2018)*10000)
```

  6. **`dplyr` review**: Look through the code above and describe what each line of code does.

The first line of code reads in the CSV with the data of the US census 2018. Then, the next line separates the state column into the state name and the dot (.) before the name. Then we select everything except the new dot column. Select allows you to see certain columns and variables. Finally, we convert the state names to all lower case through the mutate function. 

The next line of code is a left join of a data set with anothr data set. Left join is a way to combine variables and values from one data set to another one. So we are taking the starbucks_us_by_state data set and left joining it with the census data set from 2018. In order to do the left join we need to say what it should join by. So we define that by saying the variable name from the starbucks data set is equal to the one from the census. And then we mutate to add another variable of starbucks per 10,000 people which equals n divided by the estimated population of the state times 10,000. 


  7. Create a choropleth map that shows the number of Starbucks per 10,000 people on a map of the US. Use a new fill color, add points for all Starbucks in the US (except Hawaii and Alaska), add an informative title for the plot, and include a caption that says who created the plot (you!). Make a conclusion about what you observe.
  
  I observe that there are more starbucks per 10,000 people in the metropolitan areas and cities. Especially on the east coast, there are a large portion of starbucks in the New York City area. I also notice that those states that have a larger population tend to have a ton of starbucks than those states who have a smaller population comparitively. 

```{r}
#US states map information - coordinates used to draw borders
states_map <- map_data("state")

# map that colors state by number of Starbucks
starbucks_with_2018_pop_est %>%
  filter(!(state_name %in% c("alaska", "hawaii"))) %>% 
  ggplot() +
  geom_map(map = states_map,
           aes(map_id = state_name,
               fill = starbucks_per_10000)) +
  geom_point(data = Starbucks %>% 
               filter(!(`State/Province` %in% c("HI", "AK")),
                      Country == "US"),
             aes(x = Longitude, y = Latitude),
             size = .05,
             alpha = .2, 
             color = "goldenrod") +
    expand_limits(x = states_map$long, y = states_map$lat) + 
    theme_map() + 
  labs(title = "Starbucks Locations Across the US",
       fill = "Starbucks per 10,000 people",
       caption = "Plot Created by Talia C, data from Starbucks CSV") + 
  theme(legend.background = element_blank(),
        plot.title.position = "plot")
```





### A few of your favorite things (`leaflet`)

  8. In this exercise, you are going to create a single map of some of your favorite places! The end result will be one map that satisfies the criteria below. 

  * Create a data set using the `tibble()` function that has 10-15 rows of your favorite places. The columns will be the name of the location, the latitude, the longitude, and a column that indicates if it is in your top 3 favorite locations or not. For an example of how to use `tibble()`, look at the `favorite_stp_by_lisa` I created in the data R code chunk at the beginning.  

  * Create a `leaflet` map that uses circles to indicate your favorite places. Label them with the name of the place. Choose the base map you like best. Color your 3 favorite places differently than the ones that are not in your top 3 (HINT: `colorFactor()`). Add a legend that explains what the colors mean.  
  
  * Connect all your locations together with a line in a meaningful way (you may need to order them differently in the original data).  
  
  * If there are other variables you want to add that could enhance your plot, do that now.  

```{r}
# Talia's Favorite Places
favorite_places_talia <- tibble(
  place = c("Buffy's Icecream", "Chatham Seaside Links", "Chatham Light House", 
            "Nauset Beach", "Corner Store", "Yellow Umbrella Books",
            "Cape Escape Mini Golf", "Del Mar Bistro and Bar", "Chatham Anglers- Veterans' Park", 
            "Chatham Cookware"),
  long = c(-69.95572, -69.9543, -69.949856, 
           -69.93681, -70.01594, -69.95734, 
           -69.98507, -69.96943, -69.96306, -69.95784),
  lat = c(41.6795, 41.68294, 41.67133,
          41.7873, 41.70007, 41.67979, 
          41.79704, 41.68601, 41.68501, 41.68055)
  )

#create leaflet map with data 
leaflet(data = favorite_places_talia) %>% #base plot
  addTiles() %>% #base map - default is openstreet map 
  addMarkers(lng = ~long, 
             lat = ~lat, 
             label = ~place) #Adds markers - knows lat and long from names in data

#plot with lines on it 
leaflet(data = favorite_places_talia) %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>% 
  addCircles(lng = ~long, 
             lat = ~lat, 
             label = ~place, 
             weight = 10, 
             opacity = 1, 
             color = col2hex("darkred")) %>% 
  addPolylines(lng = ~long, 
               lat = ~lat, 
               color = col2hex("darkred"))
```







## Revisiting old datasets

This section will revisit some datasets we have used previously and bring in a mapping component. 

### Bicycle-Use Patterns

The data come from Washington, DC and cover the last quarter of 2014.

Two data tables are available:

- `Trips` contains records of individual rentals
- `Stations` gives the locations of the bike rental stations

Here is the code to read in the data. We do this a little differently than usualy, which is why it is included here rather than at the top of this file. To avoid repeatedly re-reading the files, start the data import chunk with `{r cache = TRUE}` rather than the usual `{r}`. This code reads in the large dataset right away.

```{r cache=TRUE}
data_site <- 
  "https://www.macalester.edu/~dshuman1/data/112/2014-Q4-Trips-History-Data.rds" 
Trips <- readRDS(gzcon(url(data_site)))
Stations<-read_csv("http://www.macalester.edu/~dshuman1/data/112/DC-Stations.csv")
```

  9. Use the latitude and longitude variables in `Stations` to make a visualization of the total number of departures from each station in the `Trips` data. Use either color or size to show the variation in number of departures. This time, plot the points on top of a map. Use any of the mapping tools you'd like.
  
```{r}
#set map 
washington_dc <- get_stamenmap(
  bbox = c(left = -77.2902, bottom = 38.7225, right = -76.7395, top = 39.0650),
  maptype = "terrain",
  zoom = 11)

#inner join to get lat and long
trips_perstation <- Trips %>% 
  group_by(sstation) %>% 
  summarize(total_departures = n()) %>% 
  inner_join(Stations, by = c("sstation" = "name"))

#graph map 
ggmap(washington_dc) +
  geom_point(data = trips_perstation, 
             aes(x = long, y = lat, color = total_departures)) +
  labs(title = "Total Departures from Different Stations Across Washington DC",
       color = " Number of Total Departures",
       caption = "Plot Created by Talia C, data from Trips") + 
  theme(plot.title.position = "plot",
        plot.caption.position = "plot")
```
  
  10. Only 14.4% of the trips in our data are carried out by casual users. Create a plot that shows which area(s) have stations with a much higher percentage of departures by casual users. What patterns do you notice? Also plot this on top of a map. I think it will be more clear what the patterns are.
  
  I notice that the there is a higher proportion of casual riders on towards the center of the cluster of points which I think is related to the fact that the center is a tourist area so many people will use bikes to get to the different tourist spots. 
  
```{r}

#set map 
washington_dc <- get_stamenmap(
  bbox = c(left = -77.2902, bottom = 38.7225, right = -76.7395, top = 39.0650),
  maptype = "terrain",
  zoom = 11)

#inner join to get lat and long
trips_perstation <- Trips %>% 
  group_by(sstation) %>% 
  summarize(total_departures = n(),
            prop_casual = mean(client == "Casual")) %>% 
  inner_join(Stations, by = c("sstation" = "name"))

#graph map 
ggmap(washington_dc) +
  geom_point(data = trips_perstation, 
             aes(x = long, y = lat, color = prop_casual)) +
  labs(title = "Total Departures from Different Stations Across Washington DC",
       color = "Proportion of Casual Riders",
       caption = "Plot Created by Talia C, data from Trips") + 
  theme(plot.title.position = "plot",
        plot.caption.position = "plot")
```
  
  
  
  
### COVID-19 data

The following exercises will use the COVID-19 data from the NYT.

  11. Create a map that colors the states by the most recent cumulative number of COVID-19 cases (remember, these data report cumulative numbers so you don't need to compute that). Describe what you see. What is the problem with this map?
  
  The problem with this map is that it does not take into account the populations of the states. It is not fair to compare the number of cases without acknowledging the differences in population. 
  
```{r}
#US states map information - coordinates used to draw borders
states_map <- map_data("state")

#data set with most recent cumulative cases of COVID-19
covid19cum <- covid19 %>% 
  group_by(state, fips) %>% 
  filter(!(state %in% c("Virgin Islands", "Northern Mariana Islands", "Puerto Rico"))) %>% 
  mutate(state = str_to_lower(state)) %>% 
  top_n(n = 1, wt = date)

# map that colors state by number of cumulative cases of COVID 19
covid19cum %>% 
  ggplot() +
  geom_map(map = states_map,
           aes(map_id = state,
               fill = cases)) +
    expand_limits(x = states_map$long, y = states_map$lat) + 
    theme_map() + 
  labs(title = "Most Recent Cumulative Number of COVID-19 Cases by State",
       fill = "",
       caption = "Plot Created by Talia C, data from COVID-19 NYT Data") + 
  theme(legend.background = element_blank(),
        plot.title.position = "plot", 
        plot.background = element_rect(fill = "lightgrey")) +
  scale_fill_continuous(labels = scales::comma)
```


  12. Now add the population of each state to the dataset and color the states by most recent cumulative cases/10,000 people. See the code for doing this with the Starbucks data. You will need to make some modifications. 
  
```{r}
#US states map information - coordinates used to draw borders
states_map <- map_data("state")

#data set with most recent cumulative cases of COVID-19
covid19cum <- covid19 %>% 
  group_by(state, fips) %>% 
  filter(!(state %in% c("Virgin Islands", "Northern Mariana Islands", "Puerto Rico"))) %>% 
  mutate(state = str_to_lower(state)) %>% 
  top_n(n = 1, wt = date)
  
# inner join Starbucks data to get Lat and Long 
covid19cum_with_2018_pop_est <- covid19cum %>% 
  left_join(census_pop_est_2018,
            by = "state") %>% 
  mutate(covid19cases_per_10000 = (cases/est_pop_2018)*10000)

# map that colors state by number of cumulative cases of COVID 19/10000
covid19cum_with_2018_pop_est %>% 
  ggplot() +
  geom_map(map = states_map,
           aes(map_id = state,
               fill = covid19cases_per_10000)) + 
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() +
  labs(title = "Covid 19 Cases Per 10,000 People Across States",
       fill = "",
       caption = "Plot Created by Talia C, data from COVID-19 NYT Data") + 
  theme(legend.background = element_blank(),
        plot.title.position = "plot", 
        plot.background = element_rect(fill = "lightgrey")) +
  scale_fill_continuous(labels = scales::comma)
```
  
  13. **CHALLENGE** Choose 4 dates spread over the time period of the data and create the same map as in exercise 12 for each of the dates. Display the four graphs together using faceting. What do you notice?
  
  With this visualization, I notice that the covid cases per 10,000 increases for each date for all the states. This is likely because the pandemic progressed as time went on and more people were testing positive for COVID. 
  
```{r}
#US states map information - coordinates used to draw borders
states_map <- map_data("state")

#data set with most recent cumulative cases of COVID-19
covid19_4dates <- covid19 %>% 
  group_by(state) %>% 
  mutate(date = ymd(date),
         state = str_to_lower(state)) %>% 
  filter(!(state %in% c("virgin islands", "northern mariana islands", "puerto rico")),
         date == "2020-04-02" | date == "2020-09-05" | date == "2020-12-18" | date == "2021-05-01")

# inner join Starbucks data to get Lat and Long 
covid19cum_with_2018_pop_est <- covid19_4dates %>% 
  left_join(census_pop_est_2018,
            by = "state") %>% 
  mutate(covid19cases_per_10000 = (cases/est_pop_2018)*10000)

# map that colors state by number of cumulative cases of COVID 19/10000
covid19cum_with_2018_pop_est %>% 
  ggplot() +
  geom_map(map = states_map,
           aes(map_id = state,
               fill = covid19cases_per_10000)) + 
  expand_limits(x = states_map$long, y = states_map$lat) +
  facet_wrap(vars(date)) +
  theme_map() +
  labs(title = "Covid 19 Cases Per 10,000 People Across States for different Dates",
       fill = "",
       caption = "Plot Created by Talia C, data from COVID-19 NYT Data") + 
  theme(legend.background = element_blank(),
        plot.title.position = "plot",
        plot.background = element_rect(fill = "lightgrey"))
```



## Minneapolis police stops

These exercises use the datasets `MplsStops` and `MplsDemo` from the `carData` library. Search for them in Help to find out more information.

  14. Use the `MplsStops` dataset to find out how many stops there were for each neighborhood and the proportion of stops that were for a suspicious vehicle or person. Sort the results from most to least number of stops. Save this as a dataset called `mpls_suspicious` and display the table.  

```{r}
#display table 
MplsStops %>% 
  group_by(neighborhood) %>% 
  summarise(n = n(),
            is_sus = problem == "suspicious",
            propsus = sum(is_sus)/n()) %>% 
  arrange(desc(n)) %>% 
  select(neighborhood, n, propsus) %>% 
  distinct()

#save new dataset 
mpls_suspicious <- MplsStops %>% 
  group_by(neighborhood) %>% 
  summarise(n = n(),
            is_sus = problem == "suspicious",
            propsus = sum(is_sus)/n()) %>% 
  arrange(desc(n)) %>% 
  distinct()
```
  
  15. Use a `leaflet` map and the `MplsStops` dataset to display each of the stops on a map as a small point. Color the points differently depending on whether they were for suspicious vehicle/person or a traffic stop (the `problem` variable). HINTS: use `addCircleMarkers`, set `stroke = FAlSE`, use `colorFactor()` to create a palette.  
  
```{r}
# Color points 
pal <- colorFactor("viridis", 
                   domain = MplsStops$problem)
#leaflet map
leaflet(MplsStops) %>% 
  addTiles() %>% 
  addCircleMarkers(fillColor = ~ pal(problem),
                   stroke = FALSE,
                   fillOpacity = 0.7,
                   weight = 0.003) %>% 
  addLegend(pal = pal, 
            values = ~ problem, 
            opacity = 0.5, 
            title = NULL,
            position = "bottomright") 
```
  
  16. Save the folder from moodle called Minneapolis_Neighborhoods into your project/repository folder for this assignment. Make sure the folder is called Minneapolis_Neighborhoods. Use the code below to read in the data and make sure to **delete the `eval=FALSE`**. Although it looks like it only links to the .sph file, you need the entire folder of files to create the `mpls_nbhd` data set. These data contain information about the geometries of the Minneapolis neighborhoods. Using the `mpls_nbhd` dataset as the base file, join the `mpls_suspicious` and `MplsDemo` datasets to it by neighborhood (careful, they are named different things in the different files). Call this new dataset `mpls_all`.

```{r}
mpls_nbhd <- st_read("Minneapolis_Neighborhoods/Minneapolis_Neighborhoods.shp", quiet = TRUE)

mpls_all <- left_join(mpls_nbhd, MplsDemo, by = c("BDNAME" = "neighborhood"))
mpls_all2 <- left_join(mpls_all, mpls_suspicious, by = c("BDNAME" = "neighborhood"))
```

  17. Use `leaflet` to create a map from the `mpls_all` data  that colors the neighborhoods by `prop_suspicious`. Display the neighborhood name as you scroll over it. Describe what you observe in the map.
  
  I observe that there is a wide range of the proportion of suspicious. However the lower proportions seem to be clustered together in the northern part of Minneapolis, while the higher amount of suspiciousness occurs in the southern part of the city. 
  
```{r}
# Color points 
pal <- colorFactor("viridis", 
                   domain = mpls_all2$propsus)
#leaflet map
leaflet(mpls_all2) %>% 
  addTiles() %>% 
  addPolygons(
    stroke = FALSE, 
    fillColor = ~pal(propsus), 
    fillOpacity = 0.7, 
    smoothFactor = 0.5, 
    highlight = highlightOptions(weight = 5, 
                                 color = "black",
                                 fillOpacity = 0.9,
                                 bringToFront = FALSE),
    label = ~ BDNAME) %>% 
  addLegend(pal = pal,
            values = ~ propsus, 
            opacity = 0.5, 
            title = NULL,
            position = "bottomright") 
```
  
  
  18. Use `leaflet` to create a map of your own choosing. Come up with a question you want to try to answer and use the map to help answer that question. Describe what your map shows. 
  
  Is there a pattern for if vehicle searches occurred or not by neighborhood? 
  
  My graph shows all the stops counted in the MplsStops dataset and if there was a vehicle search or not. As we can see from the graph, most of the time, there are mostly no searches as the graph is overwhelmingly purple. Yet sprinkled in here and there is a yellow dot which means that the vehicle was searched. I wonder if there is a correlation between the neighborhood and whether or not the vehicle was searched. 

```{r}
# Color points 
pal <- colorFactor("viridis", 
                   domain = MplsStops$vehicleSearch)
#leaflet map
leaflet(MplsStops) %>% 
  addTiles() %>% 
  addCircleMarkers(fillColor = ~ pal(vehicleSearch),
                   stroke = FALSE,
                   fillOpacity = 0.7,
                   weight = 0.003) %>% 
  addLegend(pal = pal, 
            values = ~ vehicleSearch, 
            opacity = 0.5, 
            title = NULL,
            position = "bottomright") 
```
  
  
## GitHub link

  19. Below, provide a link to your GitHub page with this set of Weekly Exercises. Specifically, if the name of the file is 04_exercises.Rmd, provide a link to the 04_exercises.md file, which is the one that will be most readable on GitHub.

[github weekly exercise #4](https://github.com/tch8/stats112_/tree/main/weekly%20exercise%20%234)

[md file link](https://github.com/tch8/stats112_/blob/main/weekly%20exercise%20%234/weekly-exercise-4-TC.md)
